### 2015-06-19 Paper_JFM/grass.tex
Final finishing touches

%\documentclass[aps,prl,reprint,twocolumn,superscriptaddress,sort&compress,10pt]{revtex4-1}  % for review and submission
% \documentclass[aps,preprint,12pt]{revtex4-1}  % for review and submission
%\documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
%\documentclass[letterpaper,10pt,prl,twocolumn,aps]{revtex4-1}
%\documentclass[aps,preprint,floatfix,prl]{revtex4-1}
\documentclass{jfm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{color}
\usepackage{longtable}
\usepackage{array}
\usepackage{amssymb}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\graphicspath{ {../Paper/images/}{../Paper/images/Growth_Rate_Cont_Plots/} }
\newcommand{\bx}{{\boldsymbol{\hat{x}}}}
\newcommand{\bn}{{\boldsymbol{\hat{n}}}}
\newcommand{\bt}{{\boldsymbol{\hat{t}}}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\grad}{\mathbf{\nabla}}
\newcommand{\del}{\partial}
\newcommand{\hg}{h_g}
%\newcommand{\Rey}{{R}}
\newcommand{\Ndg}{\tilde{N}_g}
\newcommand{\monami}{\textit{monami}}
\newcommand{\ubl}{U_\text{bl}}
\newcommand{\words}[1]{\textbf{(#1)~}}
% \newcommand{\words}[1]{{}}

\newcommand{\ReyNdg}{{\Rey\Ndg}}

\definecolor{tableShade}{gray}{0.8}
% \title{\textit{Monami} as an oscillatory hydrodynamic instability in a submerged sea grass bed}
\title{Linear stability analysis for \textit{monami} in a submerged sea grass bed}
\author{Ravi Singh}
\affiliation{Brown University, Providence RI 02912 USA}
% \author{L. Mahadevan}
% \affiliation{Harvard University, Cambridge MA 02138 USA}
\author{M. M. Bandi}
\affiliation{OIST Graduate University, Okinawa 904-0495, Japan}
\author{Amala Mahadevan}
\affiliation{Woods Hole Oceanographic Institution, Woods Hole MA 02543 USA}
\author{Shreyas Mandre}
\affiliation{Brown University, Providence RI 02912 USA}

\begin{document}
\maketitle

\begin{abstract}
The onset of \monami ~-- the synchronous waving of sea grass beds driven by a steady flow -- is modeled as a linear instability of the flow. Our model treats the drag exerted by the grass in establishing the steady flow profile, and in damping out perturbations to it. We find two distinct modes of instability, which we label Mode 1 and Mode 2. Mode 1 mechanism is closely related to Kelvin-Helmholtz instability modified to account for vegetation drag, whereas Mode 2 is unrelated to Kelvin-Helmholtz and arises from an interaction of the flow in the vegetated and unvegetated layers. The vegetation damping leads to a finite threshold flow for both these modes. The model predictions for the threshold flow criteria for onset and the resulting oscillation frequency compare well with experimental observations. The inclusion of vegetation drag differentiates our mechanism from the previous analyses of \monami.
\end{abstract}

\maketitle
\section{Introduction}
Sea grasses exhibit a rich set of dynamical behavior due to their collective interaction with fluid flows.  
The hydrodynamic processes resulting from this behavior influence many environmental processes such as sedimentation, transportation of dissolved oxygen, plant growth, and biomass production  \citep{Fonseca87,Grizzle96,Nepf99,Nepf2012}. 
One such response of the submerged grass beds to steady currents is the formation of coherent large amplitude oscillations, known as \monami ~\citep{AckermanOkubo93}.  
In this letter, we provide a hydrodynamic mechanism for the onset of these coherent oscillations.

Current explanations of \monami~\citep{Ikeda96,Ghisal02,Raupach96} invoke the existence of a shear layer at the top of the grass bed (henceforth called grass top) due to vegetation drag. 
Its instability, through a mechanism similar to the Kelvin-Helmholtz (KH) instability, is thought to lead to coherent eddies over the grass bed.
The grass responds to these eddies by deforming, which manifests itself as large amplitude synchronous oscillations.
Basic application of this model have been used to predict the observed frequency of \monami~and to understand transport in the seagrass bed \citep{Nepf00,Ghisal02,Nepf04,Okamoto12}.

However, several aspects of this explanation remain unsatisfactory. 
The instability is modeled using the inviscid Rayleigh equation and the role of drag in damping the coherent perturbations to the shear profile is ignored \citep{Raupach96}. 
It is so despite the shear layer itself being produced as a result of the vegetation drag.
Furthermore, the shear layer is assumed to be not influenced by the top and bottom boundary of the domain, although the experimentally measured thickness of the shear layer is in many cases comparable to the unvegetated water depth.
The velocity profile of the free shear layer is assumed \textit{ad hoc} to be piecewise linear \citep{Delangre06} or hyperbolic tangent \citep{Ghisal02,Raupach96}. 
While these profiles qualitatively capture the features of the profile, the dependence of the shear layer thickness on the parameters of the problem remains unexplained.
And finally, no existing theory explains the threshold flow speed, observed in the lab \citep{Ghisal02} and the field \citep{Grizzle96}, below which \monami ~is not observed.

Here we present a mathematical model for the linear instability that accounts correctly for these effects.
The vegetation is modeled as a continuum through a drag on the fluid proportional to the grass blade density. 
We solve for the steady fully-developed flow profile established due to an imposed pressure gradient in the presence of this drag, and perform linear stability analysis of this profile.
We compare the results of our calculations with data available from lab experiments and field observations.
To better understand the flow instability mechanisms and to compare them with the current explanations of \monami, we investigate the limit of asymptotically dense vegetation.

We find two modes of instability labeled Mode 1 and Mode 2.  
We find that Mode 1 resembles many characteristics of KH instability but is modified by the presence of the vegetation drag. 
We also find another mode of instability, Mode 2, unrelated to the KH instability mechanism, arising purely due to the interaction of an inviscid layer of fluid with submerged rigid vegetation.
The presence of vegetation drag leads to a finite threshold flow condition for onset of the instability for both these modes.

\section{Mathematical model}
Although \monami ~is manifest in the motion of the grass, the drag exerted by the grass bed on the flow is central to the hypothesized instability. 
The instability and the resulting flow structures persist in lab experiments even when flexible grass mimics are replaced by rigid dowels \cite{Ghisal02,Nepf06}. 
Therefore, to develop the essential mathematical model, we assume the grass blades to be rigid, on average oriented vertically.
We show that the drag of the grass results in an instability mode different from the KH instability.
\begin{figure}
\centerline{\includegraphics[scale=1]{Grass_Base_Nepf_shear}}
% \includegraphics[width=15cm]{Grass_Base_Nepf_shear}
\caption{
Schematic setup and comparison of our steady flow profile with that from the experiments in ref. \cite{Nepf04} (Case I from Table 1) %  with 1250 plants/m$^2$, plant height = 13.7$\pm 0.2$ cm and blade width of 0.64 cm)
 and its approximation with $U_0=7.28$ cm/s and $\delta = 5.02$ cm in our model. The grass extends up to $y=\hg$ in the water column of depth $2H$. 
The steady velocity profile can be decomposed into a parabolic profile in the unvegetated region, a uniform profile deep within the vegetation, and a boundary layer of thickness $\delta$ near the grass top. 
The dependence of the boundary layer thickness (estimated as $|U/U_y|$ at $y=\hg$ from the numerical solution of \eqref{base_equ}) on the vegetation density parameter $\ReyNdg$ is shown in the inset.
}
\label{basicflow}
\end{figure}
\begin{figure}
\begin{center}
\includegraphics[scale = 0.95]{new_graph_R_vs_delta}\includegraphics[scale = 0.95]{new_graph_K_vs_delta}
% \includegraphics[width=12cm]{Critical_Re_vs_delta_noshear} \\
% \vspace{-6mm} \hspace{-5mm}
% \includegraphics[width=12cm]{K_vs_shear_width_noshear}
\end{center}
\caption{
Critical Reynolds number (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density parametrized by the boundary layer thickness. 
Parameters from experiments reported in ref. \cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the top panel. 
Inset compares experimental observations of the experimentally measured dominant frequency $f_o$ (in Hz) with the predictions $f_p=\text{Im}(\sigma)$ from the solution of ~\eqref{Orr-somerfield}. 
The experimental data in the inset is obtained from refs. \cite{Ghisal02} (Ghisalberti) and \cite{Vivoni98} (Vivoni). 
In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.
}
\label{Re_vs_delta}
\end{figure}

The drag exerted by the vegetation, assumed dense, is modeled by a continuous body force $\mathbf{f}$.
This drag enters the fluid mass and momentum balance equations as 
\begin{equation}
\grad\cdot\bu = 0,\hspace{3mm} \rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad p+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
\label{navier-stokes}
\end{equation}
where $\rho$, $\bu$, and $p$ are the fluid density, velocity, and pressure respectively, $\mu$ the dynamic eddy viscosity and $\mathbf{g}$ the acceleration due to gravity. 
The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it as the form drag on the vegetation, $\mathbf{f}=N_g C_N \rho u |u| d\bx$ \cite{Nepf99,Nepf00,Nepf04}, where  $N_g$ is the blade number density per unit horizontal area, $C_{N}$ the drag coefficient, and $d$ the average blade width projected perpendicular to the flow. 
In the interest of simplicity, we model the turbulence using an eddy viscosity; see \S\ref{subsec:turbulencemodel} for a detailed discussion. 
In the field, both $C_N$, $N_g$ and $\mu$ vary with position, but we do not expect these variations to be central to the instability mechanism, and therefore take them to be constants. 
% Based on previous experiments \cite{Vivoni98,Nepf00}, we take $C_N = 1$.
% Note that only the product $C_N N_g d$ appears together in our model.

\begin{figure*}
 \includegraphics[width=\textwidth]{SetAll_imgsc}
%\end{figure}
%\begin{figure*}
%\begin{tabular}{cccc}
%{\includegraphics[height = 4cm, width = 5.8cm]{Set4_dens28_imgsc}} &
%{\includegraphics[scale = 0.67]{Set4_dens30_imgsc}} &
%{\includegraphics[height = 4cm,width = 5.8cm]{Set4_dens32_imgsc}} &
%{\includegraphics[height = 4.15cm,width=6.5cm]{Set4_dens34_imgsc}} \\
%{\includegraphics[height = 4cm, width = 5.8cm]{Set5_dens38_imgsc}} &
%{\includegraphics[scale = 0.67]{Set5_dens40_imgsc}} &
%{\includegraphics[height = 4cm, width = 5.8cm]{Set5_dens42_imgsc}} &
%{\includegraphics[height = 4.15cm,width=6.5cm]{Set5_dens46_imgsc}} \\

%{\includegraphics[scale = 0.67]{Set6_dens32_imgsc}} &
%{\includegraphics[scale = 0.67]{Set6_dens34_imgsc}} &
%{\includegraphics[scale = 0.67]{Set6_dens36_imgsc}} &
%{\includegraphics[scale = 0.67]{Set6_dens38_imgsc}}
%\end{tabular}
\caption{
$Re(\sigma)$ and the neutral curve ($Re(\sigma)$=0) as function of wavenumber and $\Rey$ for parameters shown in the corresponding panel.  
As $\Ndg$ increases, the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. 
For $\Ndg$ below (above) a critical value, Mode 1 (Mode 2) sets the threshold $\Rey$.
}
\label{K_Re_sigma_set3}
\end{figure*}

\section{Linear stability analysis}
We first calculate the fully developed steady solution $\bu = U(y)\boldsymbol{\hat{x}}$ of ~\eqref{navier-stokes} driven by constant pressure gradient $dP/dx$, satisfying
\begin{equation}
 -\frac{dP}{dx}+\mu U''(y) +S(y) \rho C_N d N_gU |U|=0
\label{base_equ}
\end{equation}
where $S(y)=1$ for $0<y<\hg$ and $S(y)=0$ for $\hg< y< 2H$. 
Eq. \eqref{base_equ} is solved subject to no shear at the boundaries, i.e., $U'(0) = U'(2H) = 0$.
The former arises for dense vegetation because the shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag~\citep{Nepf00}, whereas the latter models the free interface. 
A comparison of the steady flow profile from the solution of ~\eqref{base_equ} with experimental measurements is shown in Fig.~\ref{basicflow}.
The profile $U(y)$ has three distinct regions.
Within vegetation, it is approximately uniform with $ U(y) \approx U_g = \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$, arising from balancing the drag with pressure gradient. 
Outside the vegetation, the velocity has a simple parabolic profile. % due to the balance between viscous forces and the pressure gradient. 
At the grass top, continuity of shear stresses results in a boundary layer of thickness $\delta$. 
Denoting $\ubl$ to be the velocity scale in the boundary layer, and $U_0 = {(dP/dx)~H^2}/{\mu}$ the velocity scale in the unvegetated region, the balance between viscous forces and vegetation drag implies $(\mu \ubl/\delta^2 \sim \rho C_N d N_g \ubl^2)$, and the continuity of shear stress across the grass top implies $(\ubl/\delta \sim U_0/H)$.
Solving for $\delta$ and $\ubl$ yields $\delta/H = \ubl/U_0=(\ReyNdg)^{-1/3}$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, and $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow. 
A numerical estimate of $\delta$ ($U/U_y$ at $y=\hg$) is compared with this prediction in Fig. \ref{basicflow} (inset).
We identify the boundary layer to be analogous to the shear layer \citep{Ghisal02,Nepf04} in the previous explanation of \monami.
This dependence of $\delta$ on $N_g$ gives us a way to systematically investigate the effect of the shear layer thickness on the instability mechanism.
Fig. \ref{basicflow} also shows that the asymptotic regime of a thin boundary layer is expected to hold for $\ReyNdg \gtrsim 100$. 
In this notation, $U_g/U_0 = (\Rey \Ndg)^{-1/2}$ (used for deriving \eqref{eqn:mode2asymp}).

Next we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in ~\eqref{navier-stokes} and expand to linear order to investigate the evolution of small perturbations $(\tilde{u}, \tilde{v})$, which obey
\begin{equation}
\begin{split}
\rho(u_t+U u_x+vU_y) &= -p_x+ {\mu}\nabla^2u-2S\rho C_{N}dN_{g}Uu, \\
\rho(v_t+ Uv_x) &= -p_y+ {\mu}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0,
\end{split} \nonumber
\end{equation}
where the tilde are dropped.
These equations are non-dimensionalized using half channel height $H$, velocity $U_0$, and time $H/U_0$, leading to three non-dimensional parameters, \textit{viz.} $\Rey$, $\Ndg$, and the vegetation submergence ratio $\hg/H$. 
We also use $\delta/H$ in lieu of $\Ndg$ to parametrize the vegetation density and help elucidate the instability mechanism. 
Using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a solution of 
the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \phi(y) \right)e^{ikx+\sigma t}$ to obtain a modified Orr-Sommerfeld equation \citep{Drazin81,Chen97,Chu91} 
\begin{equation}
\begin{split}
\Rey^{-1}\left(D^2 -k^{2} \right)^2\phi &= \left[ \left({\sigma}+ikU\right) \left(D^2-k^2\right) -ikU_{yy}\right]\phi + \Ndg D\left(2 S U D \phi\right),
\label{Orr-somerfield}
\end{split}
\end{equation}
where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$. 
The growth rate $\sigma$ for a given wave number $k$ appears as an eigenvalue that allows a non-trivial solution $\phi$ of  \eqref{Orr-somerfield}.
We solve \eqref{Orr-somerfield} numerically for $\sigma$ and $\phi$.

\section{Results - unstable modes and critical parameters}
A threshold in $\Rey$, above which the flow is unstable (Re$(\sigma)>0$) for at least one $k$, emerges from the solution of ~\eqref{Orr-somerfield}. 
The dependence of this threshold $\Rey$, and the corresponding marginally stable wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown in Fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \citep{Ghisal02}.
The threshold $\Rey$ increases with the vegetation density, indicating a competition between the destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
A similar conclusion was presented for an analogous problem (flow around an emergent (i.e., $\hg>2H$) sea grass patch), but by assuming $U(y)$ to be a tanh-profile, and neglecting the viscous term \citep{White07}.
Previous calculations for terrestrial grass either exclude the vegetation drag in their models \cite{Raupach96}, or assume the mean velocity profile \textit{ad hoc} \citep{Raupach96,Delangre06}.
A threshold flow condition is not reported previously either for terrestrial or submerged marine meadows.

The frequency (Im$(\sigma)$) of the fastest growing mode agrees well with observed frequencies -- frequency of \monami, maxima in the velocity spectra, and frequency of vortex passage in lab scale experiments \citep{Ghisal02} -- for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, as shown in the inset in Fig.~\ref{Re_vs_delta}. 
The experimentally observed \monami ~wavelengths are not available for comparison.

\begin{table*}
% \footnotesize
\rowcolors{3}{tableShade}{white}  %% start alternating shades from 3rd row
\renewcommand{\arraystretch}{1.2}
 \begin{tabular}{l|c|c|c}
			& KH 				& Mode 1 		& Mode 2 \\ \hline
 Base velocity profile 	& $U(y) = U_0 \tanh(y/\delta)$			& \multicolumn{2}{c}{Equation \eqref{base_equ}} \\
 Domain 		& $-\infty < y < \infty$			& \multicolumn{2}{c}{$-1<y<1$} \\
 Inflection point	& exists at $y=0$				& \multicolumn{2}{c}{$U''(y)$ discontinuous at $y=\hg$} \\
 Shear layer thickness	& $\delta$					& \multicolumn{2}{c}{$\delta \sim  H\left(\Rey \Ndg \right)^{-1/3}$} \\
 Linearized dynamics	& Equation \eqref{eqn:Rayleigh}		& \multicolumn{2}{c}{Equation \eqref{Orr-somerfield}} \\
 Dense grass limit &  no grass included & Equation \eqref{eqn:mode1asymp} & Equation \eqref{eqn:mode2asymp}  \\
 Critical parameters	& none						& $\Rey \propto \Ndg^{2}$ 	& $\Rey \propto \Ndg$ \\
 Most unstable $k$ as $\delta \to 0$	& $\propto H/\delta$		& $\propto H/\delta$	& $O(1)$ \\
 Mode localized?	& yes, near $y=0$				& yes, near $y=\hg$			& no, spans water column
 \end{tabular}
 \caption{Comparison between KH instability and the two unstable modes resulting from solution of \ref{Orr-somerfield}.}
 \label{tab:comparison}
\end{table*}
To better understand the instability mechanism, we consider the dependence of the fastest growing wavenumber on $\delta$.
The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical $\delta$ discontinuously jumps and remains $O(1)$ (see Fig.~\ref{Re_vs_delta}). 
To aid in explaining this behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in Fig.~\ref{K_Re_sigma_set3}. 
The smallest $\Rey$ on the neutral curve (Re$(\sigma)=0$) sets the threshold. 
We observe that as $\Ndg$ increases, the unstable region splits into two; we refer to the region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''. 
The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the vegetation density increases, causing the most unstable mode to transition discontinuously.

\section{Discussion -- comparison of unstable modes with Kelvin Helmholtz}
The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ distinguishes it from the KH instability. 
\subsection{Mode 1}
We numerically observe that critical Reynolds number above which perturbation become unstable scales as  $\Rey \sim (H/\delta)^2$ (or $\Rey \propto {\Ndg}^{2}$). 
Our calculations also show that Mode 1 asymptotically localizes to the boundary layer near the grass top, and exhibits highest growth for a perturbation of  $k \sim H/\delta$ (see Fig.~\ref{Re_vs_delta}). 
The behavior of this critical Reynolds number can be understood by considering the limit $\Rey \gg 1$ and $\Ndg \gg 1$.
Estimating the size of various terms of ~\eqref{Orr-somerfield} within the boundary layer in this limit help us understand the behavior of the critical $\Rey$. 
Using $D\sim H/\delta$, $\sigma \sim O(1)$, and $U=\ubl \sim \delta/H$ in the boundary layer; the magnitude of the advection term is $\Rey (H/\delta)^2$  (or $\Rey^{5/3} \Ndg^{2/3}$), and the viscous as well as vegetation drag term are $(\delta/H)^{-4}$ (or $(\Rey \Ndg)^{4/3})$. 
The advection term, viscous term and vegetation drag terms balances when $\Rey \sim (H/\delta)^2$ (or $\Rey \sim {\Ndg}^{2}$).

To further understand the mechanism of Mode 1, we rescale \eqref{Orr-somerfield} in the the boundary layer using $\eta = y/\delta$, 
$U(y) = (\delta/H)\bar{U}(\eta)$ and $k = (H/\delta) \bar{k}$, with these scalings ~\eqref{Orr-somerfield} simplifies to
\begin{equation}
\begin{split}
\left(\bar{D}^2 -\bar{k}^{2} \right)^2\phi &= (\Rey/\Ndg^2)^{1/3} \left[ \left({\sigma}+i\bar{k}\bar{U}\right) \left(\bar{D}^2-\bar{k}^2\right) -i\bar{k}\bar{U}_{\eta\eta}\right]\phi + \bar{D}\left(2S \bar{U} \bar{D} \phi\right),
\label{eqn:mode1asymp}
\end{split}
\end{equation}
in a region of thickness O($\delta$) near $y=\hg$, where $\bar{D} = d/d\eta$. 
Since $(\Rey/\Ndg^2)$ is the only remaining parameter in \eqref{eqn:mode1asymp}, the mode shape and solution are expected to converge in the limit $\Rey \gg 1$, $\Ndg \gg 1$, but $\Rey/\Ndg^2$ fixed.
Our numerical finding confirm this expectation; the critical $\Rey$ scales as $\Ndg^2$ as shown in Fig.~\ref{Re_vs_delta} and the mode shapes are self-similar with length scale $\delta$, as shown in Fig \ref{Asymptotic_mode}.

Mode 1 shares many characteristics with the KH instability (see Table \ref{tab:comparison}). 
The fastest growing wavenumber at the critical $\Rey$ scales as $k \propto (H/\delta)$, similar to KH instability. 
The extend of the unstable mode is also localized to the boundary layer region.
Furthermore, the strong vorticity gradient $U_{yy}$ in the boundary layer plays a central role in destabilizing the flow and localizing the instability to that region.

Our detailed description of Mode 1, given by \eqref{eqn:mode1asymp} also highlights key differences with formulations of KH. 
KH is usually described using the inviscid Rayleigh's equation, 
\begin{align}
\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi, 
\label{eqn:Rayleigh}
\end{align}
and is therefore not parametrized by the Reynolds number.
Describing the instability using the Orr-Sommerfeld equation introduces the Reynolds number as a parameter, but shear flows with tanh-profiles are unstable for all values of the parameter \citep{Drazin81}.
Therefore, based on the inviscid and viscous formulations of KH instability, the origin of the threshold flow conditions observed in experiments and the field is unclear.

In our model, the vegetation, even if made very dense, may not simply be modeled by a no-slip condition.
Instead the porous nature of the vegetation implies that the flow penetrates a thin boundary layer region and establishes a strong shear profile. 
The (turbulent eddy) viscosity sets the scale of the boundary layer, and therefore for Mode 1, in this formulation.
The threshold flow condition arises from a competition between of the destabilizing role of fluid inertia, which is very similar to the one played in KH, and the vegetation drag.
The vegetation drag may not be neglected within this boundary layer, and therefore plays a central role in the Mode 1 instability mechanism.

%We expect identical asymptotic behavior for a fixed relative magnitude of the terms on the r.h.s. to those on the l.h.s. of \eqref{eqn:mode1asymp}, which is $(\Rey \delta/H)^{3/2}$ (or $\Rey/\Ndg^{1/2}$).
%Therefore the threshold obtained for Mode 1 is $\Rey \propto H/\delta$ (or $\Rey \propto \Ndg^{1/2}$) explaining the numerically observed asymptote (see Fig.~\ref{Re_vs_delta}). 
%This analysis also concludes that the mode structure is self-similar over the length scale $\delta$ for fixed $\Rey/\Ndg^{2}$; the verification of this idea is shown in Fig.~\ref{Asymptotic_mode} (inset).

\subsection{Mode 2}
The threshold condition for Mode 2 is numerically observed to be $\Rey \propto ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in Fig.~\ref{Re_vs_delta}, which can be understood by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
% Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the vegetation is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
In this limit, the non-dimensional flow in the grass bed is $U_g/U_0 \sim (\Rey \Ndg)^{-1/2} \ll 1$, and therefore $ikU \approx \sigma$ may be neglected in comparison to $\sigma$. 
Furthermore, $U_{yy}$ decays to zero within the grass outside the boundary layer . 
Outside the grass, the turbulent viscous stress term is negligible compared to the inertial term because $\Rey \gg 1$. 
Thus, \eqref{Orr-somerfield} simplifies to 
\begin{subequations}
\begin{align}
% \begin{split}
\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi,  \quad &\text{ for } y<\hg  \label{eqn:mode2asympa} \\
\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi, \quad &\text{ for } y>\hg. \label{eqn:mode2asympb}
% \end{split}
\end{align}
\label{eqn:mode2asymp}
\end{subequations}

\begin{figure}
\includegraphics[]{Asymptotic_noshear}
% \includegraphics[width=12cm]{Asymptotic_noshear}
\caption{
Plot of the neutral Mode 1 (solid) and Mode 2 (dashed) shape $|\phi|$ in the limit of small $\delta/H$ for $\hg/H=0.2$. 
% Mode 1 is shown in solid and Mode 2 is shown in dashed. The parameters for Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specified in terms of $\delta/H$) but $\Rey/\Ndg = O(1)$. 
The approach of mode shapes to each other for these small values of $\delta/H$ indicates that the dense vegetation asymptote is reached. 
Mode 1 shapes appear self-similar in shape as $\delta\to 0$.
Inset shows rescaled $|\phi|$ for Mode 1 as a function of $(y-\hg)/\delta$ approach a universal shape, indicating that an asymptotic limit has been reached. 
% The limit is not yet reached for the case $\delta/H = 0.005$ due to the influence of bottom boundary; the vegetation height in this case is comparable to the boundary layer thickness.
}
\label{Asymptotic_mode}
\end{figure}
The only remaining parameter in \eqref{eqn:mode2asymp} is $\Rey/\Ndg$. 
For fixed $\Rey/\Ndg$, the mode shape converges in the aforementioned limit, in agreement with our numerical results shown in Fig.~\ref{Asymptotic_mode}.
This convergence indicates that we have identified the correct asymptotic limit to investigate Mode 2.

The parameter $\Rey/\Ndg$ therefore sets the threshold, justifying the numerically observed asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$; see Fig.~\ref{Re_vs_delta} for comparison with numerical results).
The structure of this mode in the aforementioned limit is such that $\phi$ is continuous at $y=h_g$, but $D\phi$ undergoes a rapid transition there on the scale of boundary layer thickness $\delta$.
The eigenvalues and the mode shape are otherwise independent of $\delta$, and therefore we conclude that the boundary layer only plays a secondary role of regularizing the discontinuity in tangential velocity arising at $y=\hg$ in this instability mechanism.
The enhanced shear in the boundary layer plays no role for this mode of instability.
% We interpret Mode 2 as the instability of an inviscid fluid, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass top plays no role. 
% On the other hand, Mode 1 asymptotically localizes to the boundary layer near the grass tip, and exhibits a different asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.

Mode 2 has characteristics distinct from KH. Outside the grass, the unstable mode shape is governed by the inviscid Rayleigh's equation \eqref{eqn:Rayleigh}.
An inflection point in $U(y)$ is necessary condition for instability arising from \eqref{eqn:Rayleigh} according to Rayleigh's criteria \citep{Rayleigh1879}. 
However, for our $U(y)$ profiles, $U_{yy}(y) = -1$ above the grass and therefore does not change sign for $y>\hg$. 
Instead, the dynamics are coupled with the flow in the grass bed described by \eqref{eqn:mode2asympa} in $y< \hg$.
The absence of $U_{yy}$ in \eqref{eqn:mode2asympa} indicates that $U_{yy}$ is approximated to be zero in $y<\hg$, and therefore the positive values of $U_{yy}$ that occurs in the boundary layer do not affect this mode of instability to leading order.
Furthermore, the presence of the critical parameter $\Rey/\Ndg$ in \eqref{eqn:mode2asympa} indicates the presence of alternative destabilizing dynamics, involving the interaction of flow in the unvegetated region governed by \eqref{eqn:Rayleigh} with the flow in the vegetated region incorporating the drag.
Based on this analysis, we conclude that Mode 2 is distinct from the KH instability, and owes its existence to the vegetation drag.

% \subsection{Comparison with KH}
% Table \ref{tab:comparison} compares the two modes to each other, and to the KH instability. 
% Because the eigenfunction of Mode 1 is localized over a length scale $\delta$, it may be interpreted as the instability of the flow in the boundary layer, whereas Mode 2 may be  % understood as the instability on the scale of the water column. 
% Mode 1 appears to be superficially similar to the Kelvin Helmholtz mechanism, whereas Mode 2 arises purely from the interaction between the unvegetated water column and the flow through the vegetation. 
% The appearance of the vegetation drag parameter in the dominant balances represented by ~\eqref{eqn:mode1asymp} and ~\eqref{eqn:mode2asymp}, and the resulting threshold criteria % demonstrates its role in setting the threshold, and in distinguishing them from the KH instability.
% Vegetation drag plays a dominant role in the mechanism for both the modes, which distinguishes our analysis from the traditional KH instability.

\subsection{Role of turbulence model}\label{subsec:turbulencemodel}
We have modeled turbulence using a constant eddy viscosity. 
The simplicity of this model allows us to make progress and capture the essential features of the instability without accounting for the rich and detailed characteristics of the turbulence. 
However, this simplicity in some cases only provides a qualitatively accurate description of the flow. 
In this subsection, we present an account of the advantages and shortcomings of assuming a constant eddy viscosity to model the turbulence.

As a consequence of the constant eddy viscosity, the boundary layer thickness $\delta$ scales as $\ReyNdg^{-1/3}$.
Experimental observations show that the boundary layer thickness scales instead as $\Ndg^{-1}$ \citep{Nepf07}.
While the precise thickness of the boundary layer is governed by the details of the turbulence model, the existence of this boundary layer for dense vegetation is a general feature of the flow. 
We have captured one possible realization of this feature using constant eddy viscosity.
Experiments have also shown that a model based on mixing length $l$ better approximates the turbulent characteristics of the flow with $l = \delta$; \textit{i.e.}, the boundary layer itself establishes eddies to transport momentum. 
The eddy viscosity corresponding to this model is $\mu = \rho U \delta$, and the leading order balance between turbulent momentum transport and vegetation drag is $\mu U/\delta^2 = \rho C_N d N_g U^2$.
Substituting $\mu$ yields $\delta/H = \Ndg^{-1}$, in agreement with the experimental observations.
Within our framework, the mixing length model implies a scale for the eddy viscosity $\mu = \rho U_g \delta$ at grass top, which corresponds to an effective $\Rey = U_0H/U_g \delta$.
Furthermore, matching the slope of the velocity profile from the boundary layer to the unvegetated flow implied $U_g/U_0 \sim \delta/H$, and therefore $\Rey = (H/\delta)^2$.
Substituting this relation in $\delta/H = (\Rey \Ndg)^{-1/3}$ and solving for $\delta$ yields $\delta/H = \Ndg^{-1}$, which is in agreement with experiments and the simple scaling analysis.
This simple scaling analysis shows that the boundary layer thickness depends on the turbulence model, and indicates that turbulence models based on mixing lengths will yield more realistic scalings for boundary layer thickness.
At the same time, the qualitative features of the instability may be predicted by re-interpreting results of our analysis.

The Mode 1 instability is driven by the intense shear on the scale of the boundary layer.
The driving mechanism for this instability is similar to that of KH, and relies only on the presence of this shear as presented in the $U_{\eta\eta}$ term in equation \eqref{eqn:mode1asymp}. 
Therefore, we expect Mode 1 instability to be exhibited independent of the turbulence model. 
We further expect the fastest growing wavenumber to be proportional to $1/\delta$, and the mode to be localized to the boundary layer because these results have a basis in dimensional analysis.
The threshold parameters for Mode 1, however, may depend on the precise turbulence model used.

For the Mode 2 instability, the turbulent momentum transport is found to be irrelevant to leading order.
In the asymptotic limit of dense grass, \eqref{eqn:mode2asymp} shows that the instability is driven by the interaction of the unvegetated flow with the vegetation drag.
The precise turbulence model used influences the regularization of the sharp transition in tangential velocity across the grass top.
Therefore we expect Mode 2 to be present and its features to be preserved even if a different turbulence model is used.

\subsection{Comparison with previous models}

% Chen and Jirka (1997), White and Nepf (2007). What are their results and how do they relate to ours? These analyses have investigated instability of hyperbolic tangent profiles using modified versions of Orr Sommerfeld or Rayleigh's equations in the context of shallow water flow instabilities. They have represented the additional drag arising due to vegetation or bottom friction using a parametrization analogous to ours. 
A modified version of the Orr-Sommerfeld equation was analyzed previously by \cite{Chu91}, \cite{Chen97}, and \cite{White07} in the context of instabilities in depth-averaged shallow water flows, where bottom friction replaces or augments vegetation drag.
They assumed the steady profile to be a hyperbolic tangent, the drag to be isotropic, and the flow domain to be infinite in $y$.
\cite{White07} also neglected the eddy viscosity in their stability analysis.
While a detailed investigation needed to compare the consequence of the different assumptions is outside the scope of this paper, we discuss similarities and differences between their results and ours.
These investigations only found one unstable mode.
It is most likely so because they restricted their calculations to a parameter regime where the two modes have not yet been separated from each other, as is the case shown in Fig. \ref{K_Re_sigma_set3} for the lowest $\Ndg$.
These investigations also found that increasing the drag could further destabilize the flow, which is consistent with our interpretation of Mode 2 instability mechanism.

The analogous oscillation of terrestrial canopies in wind, known as \textit{honami} \citep{Inoue56,Raupach96}, is different because the atmospheric boundary layer is much larger than the vegetation height.
% A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically \cite{Vivoni98,Nepf00}. 
% Another difference is that terrestrial vegetation is much more rigid, whereas aquatic vegetation is buoyant \cite{Vivoni98,Ghisal02}. 
In the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represent the hydrodynamic instability for the terrestrial case.
We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large vegetation density, so as to make Mode 2 irrelevant. 
In this manner, we recover the KH-like characteristics observed in the terrestrial case.

% We now test the assumption of an undeformable grass bed due to the dominant restoring force of buoyancy, using the criteria that the buoyancy time scale be much shorter than the hydrodynamic time scale $H/U_0$.
% For a common seagrass, \textit{Zostera Marina}, the relative density difference $\Delta \rho /\rho \approx 0.25$, the volume fraction $V_f \approx 0.1$ and $H=1$ m \citep{Fonseca98}, yielding the buoyancy time scale $\sqrt{\rho H/V_f \Delta \rho g} \approx 2$ s.
% The hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore longer than the hydrodynamic time scale.
% We have neither accounted for the pre-factors appearing in the scaling argument, or considered cases when the time-scale separation is not so evident.
% Accounting for these factors  can lead to further interesting behavior \citep{Delangre06}.

While predictions of the threshold $\Rey$ for the onset of the instability and the frequency of oscillations are comparable to experimental observations, further improvement is possible.
The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
In real meadows, the drag coefficients are known to vary from bottom to tip of the grass blades \citep{Vivoni98,Nepf00}. 
The turbulence model for the flow through the meadow can also be improved from one with constant eddy viscosity \citep{Ghisal02, Nepf04}. 
Although these model improvements might lead to a better agreement between the observed and the predicted quantities, the dominance of the physical processes depicted in \eqref{eqn:mode1asymp} and \eqref{eqn:mode2asymp}, and therefore our main conclusions, are expected to remain.
\section{Conclusion}
% We now test the assumption of an undeformable grass bed due to the dominant restoring force of buoyancy, using the criteria that the buoyancy time scale be much shorter than the hydrodynamic time scale $H/U_0$.
% For a common seagrass, \textit{Zostera Marina}, the relative density difference $\Delta \rho /\rho \approx 0.25$, the volume fraction $V_f \approx 0.1$ and $H=1$ m \cite{Fonseca98}, yielding the buoyancy time scale $\sqrt{\rho H/V_f \Delta \rho g} \approx 2$ s.
% The hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore longer than the hydrodynamic time scale.
% We have neither accounted for the pre-factors appearing in the scaling argument, or considered cases when the time-scale separation is not so evident. Accounting for these factors  can lead to further interesting behavior \cite{Delangre06}.
% Indeed, the case where these time-scales are comparable can lead to interesting behavior\cite{Delangre06}, and motivates further investigation. 
In conclusion, we show that the hydrodynamic instability underlying \monami ~differs from the traditional KH due to the presence of the vegetation drag. 
The threshold flow condition observed in the field and in lab experiments arises due to the presence of this drag. 
While further investigation is needed to understand the sensitivity of the results to the various simplifying assumptions made in our model, the agreement with experiments is encouraging.
% The spatial structure of the instability modes has direct implications for transport in the grass bed; Mode 1 instability likely leads to enhanced transport near the grass tips as has been observed \cite{Nepf04,Okamoto12}, while Mode 2 instability influences the whole water column.
Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, flow through urban environments and therefore is expected to have a wider impact.

\acknowledgments
MMB was hosted by Brown University during this work and was supported by the OIST Graduate University with subsidy funding from the Cabinet Office, Japan. We thank Heidi Nepf, Marco Ghisalberti, Mitul Luhar, and L. Mahadevan for helpful discussions.

% Note the spaces between the initials
\bibliography{Grass}{}
\bibliographystyle{jfm}
%\bibliography{Grass}{}
% \bibliographystyle{plain}
%\bibliographystyle{aipauth4-1}
\end{document}